Embedding Intuitionistic into Classical Logic

### 2022-10-13 FoSSaCS/fossacs-EIICL.tex
Contributions + Related Work

A first benchmark shows that under the right circumstances this is a valid approach. Crucially, our constructions support the transfer of counter-models to validity, which is important to model checking applications, where the output of counter-examples is an appreciated feature.
\end{abstract}

@@

@@

A detailed account of the translation can be found in definition~\ref{def:fo-translation}. Finally, we benchmark our translation by translating all problems in the ILTP problem set~\cite{iltp} and applying the Vampire theorem prover~\cite{kovacs2013first} to the translated instances. While this method falls short of the absolute state of the art in general, under the right circumstances it is a reasonable approach. Since there is still a lot of room for improvement this is a hopeful first sign.

>>>>

For certain classes of benchmarks our approach can compete with state-of-the-art provers, in particular when the intuitionistic content is low. Crucially, our constructions support the transfer of counter-models to validity, which is important to model checking applications, where the output of counter-examples is an appreciated feature.
\end{abstract}

@@

\subsection*{Main Contributions}
\begin{itemize}
	\item Establishing a small model property for intuitionistic propositional logic.
	\item Embedding intuitionistic propositional logic into classical propositional logic.
	\item \resizebox*{0.96\hsize}{!}{Embedding intuitionistic propositional logic into quantified boolean formulas.}
	\item Presenting a model reduction for intuitionistic first-order logic.
	\item Embedding intuitionistic into classical first-order logic.
	\item Benchmarking our new approach to intuitionistic theorem proving.  
\end{itemize}

\subsection*{Related work}

There are various known translations between logics. Most relevant are the double negation translation~\cite{glivenko1929quelques,gentzen1936widerspruchsfreiheit}, the embedding of intuitionistic logic into S4~\cite{basicprooftheory} and of various non-classical logics into higher-order classical logic~\cite{LEO,wisniewski2016tptp}.

Of special relevance is the reduction of intuitionistic propositional logic given in~\cite{claessen2015sat} and examined in~\cite{fiorentini2019proof}, which we further expand upon.

On the other hand there have been attempts to circumvent intuitionistic logic by equipping classical proofs with operational semantics~\cite{Control1,Parigot1}, which however does not permit in general the realizability of $\exists$ and $\vee$.

There are some results relating intuitionistic and classical validity, most famously for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}, but also in some other cases~\cite{schwichtenberg}. However, they only apply in very limited settings.

Finally, of course dedicated theorem provers for intuitionistic logic exists, generally using the inverse method~\cite{mclaughlin2008imogen,pavlov2017whaleprover}, but more recently also based on connection-based calculi~\cite{otten2008leancop,otten2021nanocop}, which also work well for classical logic.

@@

A detailed account of the translation can be found in definition~\ref{def:fo-translation}. Finally, we benchmark our translation by translating all problems in the ILTP problem set~\cite{iltp} and applying the Vampire theorem prover~\cite{kovacs2013first} to the translated instances. While this method falls short of the absolute state-of-the-art in general, for certain subsets of benchmarks it beats existing provers, in particular when the intuitionistic content is low. Since there is still a lot of room for improvement this is a hopeful first sign.
